---
title: "Constraints on Caugi Graphs"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Constraints on Caugi Graphs}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(caugi)
```

## Overview

`caugi_constraints()` lets you express structural constraints on a graph
as Boolean formulas over caugi's existing query predicates. The same
object covers background knowledge for causal discovery, sensitivity
analysis ("would this edge flip still satisfy?"), assumption stress
tests, and composing expert priors with algorithmic output.

This vignette is a quick tour of the surface. The full design is in
`extras/design/constraints-plan.md` in the repository.

> The constraint system is **experimental**. The surface is stable
> enough to build against, but the native solver / encoder (Phase 2 in
> the design doc) is not yet wired up.

## A worked example

```{r}
cg <- caugi(
  A %-->% B + C,
  B %-->% D,
  C %-->% D,
  class = "DAG"
)

ctr <- caugi_constraints(
  A %-->% B,                       # this edge must hold
  !(D %in% ancestors(A)),          # D must not reach A
  c("A") %<<% c("B", "C") %<<% c("D")   # topological precedence
)

satisfies(cg, ctr)
```

When something fails, `violations()` tells you which formula:

```{r}
bad <- caugi_constraints(
  A %-->% B,
  B %-->% A          # this one will fail
)
violations(cg, bad)
```

## Atom vocabulary

Atoms are the leaves of the AST. Every form composes with the boolean
combinators below.

### Edge atoms

Any registered edge operator: `%-->%`, `%---%`, `%<->%`, `%o->%`,
`%--o%`, `%o-o%`.

```{r}
caugi_constraints(A %-->% B, A %---% B)
```

### Set-membership

R's built-in `%in%` over caugi's query functions. Tier-A queries
(`parents`, `children`, `neighbors`, `spouses`, `markov_blanket`,
`districts`) and tier-B queries (`ancestors`, `descendants`,
`anteriors`, `posteriors`) are recognised.

```{r}
caugi_constraints(
  A %in% parents(Y),
  B %in% ancestors(Y),
  C %in% markov_blanket(Y)
)
```

### Topological precedence

`L %<<% R` desugars to a conjunction of `!(r %in% ancestors(l))` over
the cartesian product. Chains expand into adjacent-pair desugarings.

```{r}
caugi_constraints(
  c("A") %<<% c("B", "C") %<<% c("D")
)
```

### Predicates

```{r}
caugi_constraints(
  acyclic(),
  collider(A, B, C),     # B is a collider on the triple (A, B, C)
  v_structure(A, B, C),  # collider + A, C non-adjacent
  dsep(X, Y, Z)          # d-separation (evaluator-only)
)
```

## Boolean structure

Inside `caugi_constraints(...)`, use `!`, `&` / `&&`, `|` / `||`, plus
`xor()` and `implies()`.

```{r}
caugi_constraints(
  xor(A %-->% B, B %-->% A),
  implies(A %in% ancestors(Y), !(Y %-->% A))
)
```

## Quantifiers

`forall(X, body)` and `exists(Z, body)`. The variable ranges over the
nodes currently in the graph; multi-variable forms use `c(...)`.

```{r}
caugi_constraints(
  forall(X, !(X %in% ancestors(X))),         # no node is its own ancestor
  exists(Z, Z %-->% B & Z %-->% C)           # some node points at both B and C
)
```

## Cardinality

`at_most(k, set)`, `at_least(k, set)`, `exactly(k, set)`. The `set` is
either a `c(...)` of formulas or a recognised query call.

```{r}
caugi_constraints(
  forall(X, at_most(3, parents(X))),         # bounded in-degree
  at_least(1, c(A %-->% Y, B %-->% Y))       # at least one of these edges
)
```

## Composing constraint sets

`caugi_constraints` instances support boolean composition between
sets. `ctr1 & ctr2` concatenates the formula lists; `ctr1 | ctr2`
produces a single disjunction.

```{r}
expert  <- caugi_constraints(A %-->% B)
algo    <- caugi_constraints(B %-->% D)
combined <- expert & algo
satisfies(cg, combined)
```

You can also attach a constraint set to a graph for later retrieval:

```{r}
cg2 <- with_constraints(cg, combined)
satisfies(cg2, constraints(cg2))
```

## User-defined predicates

`caugi_predicate(fn)` lets you name a reusable constraint pattern.
Inside `caugi_constraints(...)`, an invocation of a named predicate is
inlined — its body is substituted with the supplied arguments and
classified through the normal pipeline. Predicates can call other
predicates.

```{r}
no_unmeasured_confounder <- caugi_predicate(function(X, Y) {
  forall(Z, implies(
    (Z %in% ancestors(X)) & (Z %in% ancestors(Y)),
    observed(Z)
  ))
})

instrument <- caugi_predicate(function(Z, X, Y) {
  (Z %in% ancestors(X)) & !(Z %in% ancestors(Y))
})

ctr <- caugi_constraints(
  no_unmeasured_confounder("A", "D"),
  instrument("Z", "A", "D")
)
```

The named predicates carry no causal-discovery vocabulary into caugi
itself — downstream packages (e.g. causalDisco) can ship a library of
patterns of their own.

## Solver-backed operations

`consistent()` and `enumerate()` go through the native splr backend
rather than the R-side evaluator. `consistent()` answers whether
**any** graph of the chosen class satisfies the constraints;
`enumerate()` returns up to `limit` such graphs.

Solver classes supported: `"DAG"`, `"UG"`, `"PDAG"`, `"ADMG"`.

```{r}
# Is there any DAG over {A, B, C} with A --> B and forbidden B --> A?
consistent(
  caugi_constraints(A %-->% B, !(B %-->% A)),
  nodes = c("A", "B", "C"),
  class = "DAG"
)

# All DAGs on 3 nodes — the Robinson sequence gives 25.
length(enumerate(caugi_constraints(), c("A", "B", "C"), class = "DAG", limit = 100))

# All UGs on 3 nodes — 2^C(3,2) = 8.
length(enumerate(caugi_constraints(), c("A", "B", "C"), class = "UG", limit = 50))

# ADMG with a latent-confounder pattern.
ctr <- caugi_constraints(
  A %-->% Y,
  U %<->% A,
  U %<->% Y
)
consistent(ctr, c("A", "U", "Y"), class = "ADMG")
```

`enumerate()` is exact: it returns every distinct graph up to the
requested limit, with each result a `caugi` object that
`satisfies(cg, ctr)` confirms holds.

Each class has its own edge vocabulary and structural invariants:

- **DAG**: `-->` only; acyclic.
- **UG**: `---` only.
- **PDAG**: `-->` and `---`; at most one per pair; acyclic on the
  directed substructure.
- **ADMG**: `-->` and `<->`; at most one per pair; acyclic on the
  directed substructure (bidirected edges don't participate in
  cycles).

Membership queries route correctly per class — `B %in% spouses(A)`
is always false in a DAG / UG / PDAG, and routes to the bidirected
edge variable in ADMG. `parents` and `children` only make sense for
classes that allow `-->`.

## What's not yet built

- `MPDAG`, `AG`, `MAG`, `PAG` classes in the solver.
- `observed()`, `connected()`, `markov_blanket()` membership, and
  `anteriors`/`posteriors` reach as constraints.
- `entails(ctr, atom)` — checking whether a constraint set forces a
  particular atom in every model.
- `dsep()` in constraint position (evaluator-only — encoder rejects).
- Serialization (`read_constraints` / `write_constraints`).

The surface above is stable; the unimplemented pieces will arrive in
subsequent caugi releases.
