---
title: "Performance"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Performance}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r}
#| label: setup
#| include: false

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.align = "center",
  fig.width = 7,
  warning = FALSE,
  message = FALSE,
  echo = FALSE
)
```

```{r}
#| label: data
#| include: false

benchmarks <- readRDS("data/benchmarks.rds")
metadata <- jsonlite::fromJSON("data/metadata.json", simplifyVector = FALSE)
benchmarks$median_ms <- benchmarks$median_ns / 1e6
```

This article compares the performance of `caugi` against R packages
([`igraph`](https://r.igraph.org/), [`bnlearn`](https://www.bnlearn.com/),
[`dagitty`](https://dagitty.net/),
[`ggm`](https://cran.r-project.org/package=ggm),
[`pcalg`](https://pcalg.r-forge.r-project.org/)), the Python package
[`pgmpy`](https://pgmpy.org/), and the Java library
[Tetrad](https://www.cmu.edu/dietrich/philosophy/tetrad/). We focus on the
practical trade-offs that arise from different core data structures and design
choices.

The headline is: `caugi` frontloads computation. That is, `caugi` spends more
effort when constructing a graph and preparing indexes, so that queries are
wicked fast. The high performance is also due to the Rust backend.

Benchmarks below are **precomputed**, not run when this article is rendered.
Last run on `r metadata$generated` on host `r metadata$host$nodename`
(`r metadata$host$sysname`).

Versions:

- caugi `r metadata$versions$caugi`
- igraph `r metadata$versions$igraph`
- bnlearn `r metadata$versions$bnlearn`
- dagitty `r metadata$versions$dagitty`
- ggm `r metadata$versions$ggm`
- pcalg `r metadata$versions$pcalg`
- pgmpy `r metadata$versions$pgmpy`
- Tetrad `r metadata$versions$tetrad`.

### Design choices

#### Compressed Sparse Row (CSR) Representation

The core data structure in `caugi` is a compressed sparse row (CSR)
representation of the graph. CSR representations store, for each vertex, a
contiguous slice of neighbor IDs with a pointer (offset) array that marks the
start and end of each slice. This format is memory-efficient for sparse graphs.
The `caugi` graph object also stores important query information in the object,
leading to parent, child, and neighbor queries to be done in $\mathcal{O}(1)$
time. This yields a larger memory footprint, but the trade-off is that queries
are extremely fast.

#### Mutation and Lazy Building

The `caugi` graph objects are expensive to build. This is the performance
downside of using `caugi`. For each time we make a modification to a `caugi`
graph object, we need to rebuild the graph completely since the graph object is
immutable by design. This has complexity $\mathcal{O}(|V| + |E|)$, where $V$ is
the vertex set and $E$ is the edge set.

However, the graph object will only be rebuilt when the user either calls
`build()` directly or queries the graph. Therefore, you do not need to worry
about wasting compute time by iteratively making changes to a `caugi` graph
object, as the graph rebuilds lazily when queried. By doing this, `caugi` graphs
*feel* mutable, but, in reality, they are not.

By doing it this way, we ensure

- that the graph object is always in a consistent state when queried and
- that queries are as fast as possible,

while keeping the user experience smooth.

### Comparison

The comparison covers six operations across a parameter grid of graph sizes $n$
and edge densities. All packages run on the same generated DAG fixtures so query
results are directly comparable.

```{r}
#| label: helpers
#| include: false

library(ggplot2)

pkg_levels <- c(
  "caugi",
  "igraph",
  "bnlearn",
  "dagitty",
  "ggm",
  "pcalg",
  "pgmpy",
  "tetrad"
)
benchmarks$package <- factor(benchmarks$package, levels = pkg_levels)

plot_op <- function(op) {
  d <- benchmarks[benchmarks$operation == op, ]
  ggplot(d, aes(n, median_ms, color = package)) +
    geom_line() +
    geom_point() +
    scale_x_log10() +
    scale_y_log10() +
    facet_wrap(
      ~p_raw,
      labeller = as_labeller(c("0.5" = "Sparser (p_raw = 0.5)", "0.9" = "Denser (p_raw = 0.9)"))
    ) +
    labs(
      x = "Number of nodes",
      y = "Median time (ms)",
      color = NULL,
      title = paste("Operation:", op)
    ) +
    theme_minimal(base_size = 11) +
    theme(legend.position = "bottom")
}
```

#### Relational Queries

Direct neighbor lookups (`parents`, `children`) are $O(1)$ for `caugi`,
`igraph`, `bnlearn`, and Tetrad---they all maintain adjacency tables. `dagitty`,
`ggm`, and `pcalg` re-derive neighbors from a string DSL or adjacency matrix on
each call and pay a steep linear-in-edges cost; `pgmpy` is a thin wrapper over
NetworkX's predecessor/successor lookups.

```{r}
#| label: plot-parents

plot_op("parents")
```

```{r}
#| label: plot-children

plot_op("children")
```

#### Ancestors and Descendants

These require transitive closure (BFS/DFS over the directed graph). Here
`caugi`'s CSR pays off---both forward and reverse adjacency are precomputed, so
traversal is tight pointer-chasing in Rust.

```{r}
#| label: plot-ancestors

plot_op("ancestors")
```

```{r}
#| label: plot-descendants

plot_op("descendants")
```

#### D-separation

For each fixture we compute a valid backdoor adjustment set with
`caugi::adjustment_set()` once, and pass that same conditioning set to every
package's d-separation routine. Tetrad uses the m-separation generalisation
(equivalent on DAGs). `pcalg::dsep()` follows Lauritzen's moralization-based
test on a `graphNEL`.

```{r}
#| label: plot-dsep

plot_op("d_separated")
```

#### Subgraph Extraction

Subgraph extraction is where we explicitly test graph **building** performance.
For `caugi` we time `subgraph(cg, nodes) |> build()` together so the CSR rebuild
cost is captured. `igraph`, `bnlearn`, `pgmpy`, and Tetrad all return
adjacency-backed subgraphs without an analogous index rebuild step. `pcalg`
exposes no native subgraph constructor (users typically call
`graph::subGraph()` on the underlying `graphNEL`), so it is omitted here.

```{r}
#| label: plot-subgraph

plot_op("subgraph")
```

### Session Info

```{r}
#| label: session-info

sessionInfo()
```
