---
title: "Performance"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Performance}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r}
#| label: setup
#| include: false
#| purl: false

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.align = "center",
  fig.width = 7,
  fig.height = 4,
  warning = FALSE,
  message = FALSE,
  echo = FALSE
)
```

```{r}
#| label: bench-setup
#| include: false
#| eval: false

suppressPackageStartupMessages({
  library(caugi)
  library(bench)
  library(data.table)
  library(jsonlite)
  library(igraph)
  library(bnlearn)
  library(dagitty)
  library(ggm)
  library(pcalg)
  library(graph)
})

FIXTURES_DIR <- "fixtures"
RESULTS_DIR <- "results"
dir.create(RESULTS_DIR, showWarnings = FALSE, recursive = TRUE)

spec <- jsonlite::fromJSON(
  file.path(FIXTURES_DIR, "spec.json"),
  simplifyVector = FALSE
)
.skip_table <- if (is.null(spec$skip)) list() else spec$skip
```

```{r}
#| label: bench-load-fixture
#| include: false
#| eval: false

load_fixture <- function(fx) {
  edges_path <- file.path(FIXTURES_DIR, fx$edges_file)
  edges <- data.table::fread(
    edges_path,
    col.names = c("from", "to"),
    colClasses = "character"
  )
  nodes <- paste0("V", seq_len(fx$n))

  cg <- caugi::add_edges(
    caugi::caugi(class = "DAG", nodes = nodes),
    from = edges$from,
    edge = rep("-->", nrow(edges)),
    to = edges$to
  )
  cg <- caugi::build(cg)

  ig <- igraph::graph_from_data_frame(
    as.data.frame(edges),
    directed = TRUE,
    vertices = data.frame(name = nodes)
  )

  am <- as.matrix(igraph::as_adjacency_matrix(ig))
  storage.mode(am) <- "integer"

  bn <- bnlearn::empty.graph(nodes)
  bnlearn::amat(bn) <- am

  dg_str <- paste0(
    "dag {\n",
    paste(sprintf("%s -> %s", edges$from, edges$to), collapse = "\n"),
    "\n}"
  )
  dg <- dagitty::dagitty(dg_str)

  # pcalg's searchAM expects the PAG-style 0/1/2/3 amat coding (mark at the
  # column-end of edge {row, col}): for a -> b, amat[a,b] = 2 (arrowhead at b)
  # and amat[b,a] = 3 (tail at a).
  amat_pag <- matrix(0L, nrow(am), ncol(am), dimnames = dimnames(am))
  amat_pag[am == 1L] <- 2L
  amat_pag[t(am) == 1L] <- 3L

  # pcalg::dsep operates on a graphNEL.
  gNEL <- igraph::as_graphnel(ig)

  list(
    cg = cg,
    ig = ig,
    bn = bn,
    dg = dg,
    am = am,
    amat_pag = amat_pag,
    gNEL = gNEL,
    nodes = nodes
  )
}
```

```{r}
#| label: bench-helpers
#| include: false
#| eval: false

# bench::mark returns a tibble; we convert each row into a long-format record.
mark_to_rows <- function(bm, operation, fx) {
  pkg <- as.character(bm$expression)
  data.table::data.table(
    language = rep("R", length(pkg)),
    package = pkg,
    operation = rep(operation, length(pkg)),
    fixture_id = rep(fx$id, length(pkg)),
    n = rep(fx$n, length(pkg)),
    p = rep(fx$p, length(pkg)),
    n_edges = rep(fx$n_edges, length(pkg)),
    median_ns = as.numeric(bm$median) * 1e9,
    min_ns = as.numeric(bm$min) * 1e9,
    total_time_ns = as.numeric(bm$total_time) * 1e9,
    n_iter = as.integer(bm$n_itr),
    mem_alloc_bytes = as.numeric(bm$mem_alloc)
  )
}

# A skip rule matches when its package + operation match (with "*" as
# wildcard) and the fixture's n / id fall within optional bounds.
pkg_skip <- function(pkg, op, fx, skip_table = .skip_table) {
  if (length(skip_table) == 0L) {
    return(FALSE)
  }
  for (rule in skip_table) {
    pkg_match <- identical(rule$package, pkg) || identical(rule$package, "*")
    op_match <- identical(rule$operation, op) || identical(rule$operation, "*")
    n_min_ok <- is.null(rule$n_min) || fx$n >= rule$n_min
    n_max_ok <- is.null(rule$n_max) || fx$n <= rule$n_max
    id_ok <- is.null(rule$fixture_id) || identical(rule$fixture_id, fx$id)
    if (pkg_match && op_match && n_min_ok && n_max_ok && id_ok) {
      return(TRUE)
    }
  }
  FALSE
}

# Filter the per-package expression list against the skip table, then call
# bench::mark on the survivors and serialise the result.
run_bench <- function(calls, operation, fx) {
  keep <- !vapply(
    names(calls),
    pkg_skip,
    logical(1),
    op = operation,
    fx = fx
  )
  if (!any(keep)) {
    return(NULL)
  }
  args <- c(
    calls[keep],
    list(check = FALSE, min_iterations = 5L, time_unit = "s")
  )
  bm <- eval(as.call(c(quote(bench::mark), args)))
  mark_to_rows(bm, operation, fx)
}
```

```{r}
#| label: data
#| include: false
#| purl: false

benchmarks <- readRDS("data/benchmarks.rds")
metadata <- jsonlite::fromJSON("data/metadata.json", simplifyVector = FALSE)
benchmarks$median_ms <- benchmarks$median_ns / 1e6
```

This vignette compares the performance of `caugi` against R packages
([`igraph`](https://r.igraph.org/), [`bnlearn`](https://www.bnlearn.com/),
[`dagitty`](https://dagitty.net/),
[`ggm`](https://cran.r-project.org/package=ggm),
[`pcalg`](https://pcalg.r-forge.r-project.org/)), the Python package
[`pgmpy`](https://pgmpy.org/), and the Java library
[Tetrad](https://www.cmu.edu/dietrich/philosophy/tetrad/). We focus on the
practical trade-offs that arise from different core data structures and design
choices.

The headline is: `caugi` frontloads computation. That is, `caugi` spends more
effort when constructing a graph and preparing indexes, so that queries are
wicked fast. The high performance is also due to the Rust backend.

A note on Tetrad: it is included as a reference point—the de facto gold
standard implementation in causal discovery—but the comparison is not strictly
apples-to-apples, since Tetrad runs on the JVM (compiled Java) while the R
packages and `caugi` run with R-level dispatch overhead on every call.

Benchmarks below are **precomputed**, with the last run on
`r metadata$generated` on host `r metadata$host$nodename`
(`r metadata$host$sysname`).

Versions:

- caugi `r metadata$versions$caugi`
- igraph `r metadata$versions$igraph`
- bnlearn `r metadata$versions$bnlearn`
- dagitty `r metadata$versions$dagitty`
- ggm `r metadata$versions$ggm`
- pcalg `r metadata$versions$pcalg`
- pgmpy `r metadata$versions$pgmpy`
- Tetrad `r metadata$versions$tetrad`.

## Design Choices

### Compressed Sparse Row (CSR) Representation

The core data structure in `caugi` is a compressed sparse row (CSR)
representation of the graph. CSR representations store, for each vertex, a
contiguous slice of neighbor IDs with a pointer (offset) array that marks the
start and end of each slice. This format is memory-efficient for sparse graphs.
The `caugi` graph object also stores important query information directly,
allowing parent, child, and neighbor queries to be done in $\mathcal{O}(1)$
time. This yields a larger memory footprint, but the trade-off is that queries
are extremely fast.

### Mutation and Lazy Building

The `caugi` graph objects are expensive to build. This is the performance
downside of using `caugi`. Each time we make a modification to a `caugi`
graph object, we need to rebuild the graph completely since the graph object is
immutable by design. This has complexity $\mathcal{O}(|V| + |E|)$, where $V$ is
the vertex set and $E$ is the edge set.

However, the graph object will only be rebuilt when the user either calls
`build()` directly or queries the graph. Therefore, you do not need to worry
about wasting compute time by iteratively making changes to a `caugi` graph
object, as the graph rebuilds lazily when queried. By doing this, `caugi` graphs
*feel* mutable, but, in reality, they are not.

By doing it this way, we ensure

- that the graph object is always in a consistent state when queried and
- that queries are as fast as possible,

while keeping the user experience smooth.

## Comparison

The comparison covers seven operations across a parameter grid of graph sizes
$n$ and edge densities. All packages run on the same generated DAG fixtures so
query results are directly comparable.

The grid covers $n \in \{100, 1\,000, 10\,000\}$ at two density levels,
parametrized by the **average in-degree** $d$ (equivalently the average
out-degree, or total edges divided by number of nodes). We sample DAGs with
per-edge probability $p = 2d / (n - 1)$, which makes the expected in-degree
exactly $d$ regardless of $n$. The grid uses $d = 3$ for the sparser cells and
$d = 6$ for the denser ones, so a node has on average $\approx 3$ (or
$\approx 6$) parents and $\approx 3$ (or $\approx 6$) children, with edge counts
that grow linearly with $n$. The slowest packages (`dagitty`, `ggm`, plus
`bnlearn::dsep` and `pcalg::dsep`) are skipped at $n = 10\,000$ via the `skip`
table in `spec.json`; the affected lines simply end at $n = 1\,000$.

```{r}
#| label: helpers
#| include: false
#| purl: false

library(ggplot2)

pkg_levels <- c(
  "caugi",
  "igraph",
  "bnlearn",
  "dagitty",
  "ggm",
  "pcalg",
  "pgmpy",
  "tetrad"
)
benchmarks$package <- factor(benchmarks$package, levels = pkg_levels)

# Stable package → colour mapping. Used in every plot so a package keeps the
# same colour even when it is missing from a panel (e.g. dagitty at n=10000).
# ColorBrewer Set1 + near-black for Tetrad.
pkg_colors <- c(
  caugi = "#E41A1C",
  igraph = "#377EB8",
  bnlearn = "#4DAF4A",
  dagitty = "#984EA3",
  ggm = "#FF7F00",
  pcalg = "#A65628",
  pgmpy = "#F781BF",
  tetrad = "#1B1B1B"
)

plot_op <- function(op) {
  d <- benchmarks[benchmarks$operation == op, ]
  op_manual <- switch(op, d_separated = "d-separation", op)
  op_title <- tools::toTitleCase(gsub("_", " ", op_manual))

  ggplot(d, aes(n, median_ms, color = package)) +
    geom_line() +
    geom_point() +
    scale_x_log10() +
    scale_y_log10() +
    scale_color_manual(values = pkg_colors, drop = FALSE) +
    facet_wrap(
      ~avg_degree,
      labeller = as_labeller(c(
        "3" = "Sparser (avg degree 3)",
        "6" = "Denser (avg degree 6)"
      ))
    ) +
    labs(
      x = "Number of nodes",
      y = "Median time (ms)",
      color = NULL,
      title = op_title
    ) +
    theme_minimal(base_size = 11) +
    theme(legend.position = "bottom")
}
```

### Relational Queries

Direct neighbor lookups (`parents`, `children`) are $O(1)$ for `caugi`,
`igraph`, `bnlearn`, and Tetrad---they all maintain adjacency tables. `pgmpy` is
a thin wrapper over NetworkX's predecessor/successor lookups. `pcalg` and `ggm`
scan a row or column of the adjacency matrix on each call ($O(n)$), with `ggm`
paying noticeably more R-level overhead. `dagitty` is the outlier: it re-derives
neighbors from its string DSL representation each call and pays a
linear-in-edges cost.

```{r}
#| label: bench-parents
#| eval: false
#| echo: true

bench_parents <- function(graphs, fx) {
  v <- fx$test_node
  v_idx <- match(v, graphs$nodes)

  calls <- list(
    caugi = bquote(caugi::parents(graphs$cg, .(v))),
    igraph = bquote(igraph::neighbors(graphs$ig, .(v), mode = "in")),
    bnlearn = bquote(bnlearn::parents(graphs$bn, .(v))),
    dagitty = bquote(dagitty::parents(graphs$dg, .(v))),
    ggm = bquote(ggm::pa(.(v), graphs$am)),
    pcalg = bquote(pcalg::searchAM(graphs$amat_pag, .(v_idx), type = "pa"))
  )

  run_bench(calls, "parents", fx)
}
```

```{r}
#| label: plot-parents
#| purl: false

plot_op("parents")
```

```{r}
#| label: bench-children
#| eval: false
#| echo: true

bench_children <- function(graphs, fx) {
  v <- fx$test_node
  v_idx <- match(v, graphs$nodes)

  calls <- list(
    caugi = bquote(caugi::children(graphs$cg, .(v))),
    igraph = bquote(igraph::neighbors(graphs$ig, .(v), mode = "out")),
    bnlearn = bquote(bnlearn::children(graphs$bn, .(v))),
    dagitty = bquote(dagitty::children(graphs$dg, .(v))),
    ggm = bquote(ggm::ch(.(v), graphs$am)),
    pcalg = bquote(pcalg::searchAM(graphs$amat_pag, .(v_idx), type = "ch"))
  )

  run_bench(calls, "children", fx)
}
```

```{r}
#| label: plot-children
#| purl: false

plot_op("children")
```

As you can see from the plots, Tetrad is fastest (sub-microsecond), with `pgmpy`
and `caugi` close behind; all three stay flat as $n$ grows, as do `igraph` and
`bnlearn`. The cost of `pcalg`'s $O(n)$ row scan, hidden at small $n$, becomes
visible by $n = 10000$, where it climbs to a few tenths of a millisecond. `ggm`
and `dagitty` pay the most for re-deriving neighbors on every call: by
$n = 1000$ they are already one to two orders of magnitude slower than the rest,
and both are skipped at $n = 10000$ (their lines simply stop). For the
front-runners the absolute times are tiny and the gaps between them mostly
reflect dispatch overhead differences between Python, R, and Java rather than
anything that matters in practice.

### Ancestors and Descendants

These require transitive closure (BFS/DFS over the directed graph). Here
`caugi`'s CSR pays off---both forward and reverse adjacency are precomputed, so
traversal is tight pointer-chasing in Rust.

```{r}
#| label: bench-ancestors
#| eval: false
#| echo: true

bench_ancestors <- function(graphs, fx) {
  v <- fx$test_node
  v_idx <- match(v, graphs$nodes)

  calls <- list(
    caugi = bquote(caugi::ancestors(graphs$cg, .(v))),
    igraph = bquote(igraph::subcomponent(graphs$ig, .(v), mode = "in")),
    bnlearn = bquote(bnlearn::ancestors(graphs$bn, .(v))),
    dagitty = bquote(dagitty::ancestors(graphs$dg, .(v))),
    pcalg = bquote(pcalg::searchAM(graphs$amat_pag, .(v_idx), type = "an"))
  )

  run_bench(calls, "ancestors", fx)
}
```

```{r}
#| label: plot-ancestors
#| purl: false

plot_op("ancestors")
```

```{r}
#| label: bench-descendants
#| eval: false
#| echo: true

bench_descendants <- function(graphs, fx) {
  v <- fx$test_node
  v_idx <- match(v, graphs$nodes)

  calls <- list(
    caugi = bquote(caugi::descendants(graphs$cg, .(v))),
    igraph = bquote(igraph::subcomponent(graphs$ig, .(v), mode = "out")),
    bnlearn = bquote(bnlearn::descendants(graphs$bn, .(v))),
    dagitty = bquote(dagitty::descendants(graphs$dg, .(v))),
    pcalg = bquote(pcalg::searchAM(graphs$amat_pag, .(v_idx), type = "de"))
  )

  run_bench(calls, "descendants", fx)
}
```

```{r}
#| label: plot-descendants
#| purl: false

plot_op("descendants")
```

`caugi` is the most consistent here: sub-millisecond on every fixture for both
ancestors and descendants, and barely moving as $n$ grows.
`igraph::subcomponent()` and `pgmpy`'s NetworkX-backed traversal also stay fast
across the grid. Descendants on the dense $n = 10000$ fixture is where the
design differences blow up: Tetrad takes about ten *seconds*, while `pcalg` and
`bnlearn` reach several hundred milliseconds---`caugi` finishes the same query
in roughly a tenth of a millisecond, a five-orders-of-magnitude gap to Tetrad.
`dagitty` is again one to two orders of magnitude slower than the front-runners
on the sizes it can handle, paying its DSL-reparsing cost on every call, and is
skipped at $n = 10000$.

### Markov Blanket

The Markov blanket of a node $v$ is
`parents(v) ∪ children(v) ∪ parents(children(v))`. `caugi`, `bnlearn`, and
Tetrad have direct accessors backed by adjacency tables; `dagitty` re-derives it
from its DSL representation each call; `pgmpy` walks NetworkX edges. `pcalg` and
`ggm` do not expose a Markov-blanket helper and are omitted.

```{r}
#| label: bench-markov-blanket
#| eval: false
#| echo: true

bench_markov_blanket <- function(graphs, fx) {
  v <- fx$test_node

  calls <- list(
    caugi = bquote(caugi::markov_blanket(graphs$cg, .(v))),
    bnlearn = bquote(bnlearn::mb(graphs$bn, .(v))),
    dagitty = bquote(dagitty::markovBlanket(graphs$dg, .(v)))
  )

  run_bench(calls, "markov_blanket", fx)
}
```

```{r}
#| label: plot-markov-blanket
#| purl: false

plot_op("markov_blanket")
```

`caugi`, `pgmpy`, and `bnlearn` all resolve Markov blankets in microseconds
across the entire grid, and Tetrad stays well under a millisecond too. `dagitty`
is the outlier, climbing into the hundreds of milliseconds on the dense
$n = 1000$ fixture because it re-derives the blanket from its DSL representation
on every call; it is skipped at $n = 10000$.

### D-Separation

For each fixture we pick a random `(X, Y)` pair and compute a **minimal
d-separator** `Z` with `caugi::minimal_separator()`---that is, a smallest set of
nodes whose conditioning blocks every path between `X` and `Y` in the DAG. The
same triple `(X, Y, Z)` is then passed to every package's d-separation routine,
so every package answers "is `X ⫫ Y | Z`?" on identical inputs. Tetrad uses the
m-separation generalisation (equivalent on DAGs). `pcalg::dsep()` follows
Lauritzen's moralization-based test on a `graphNEL`.

```{r}
#| label: bench-dsep
#| eval: false
#| echo: true

bench_dsep <- function(graphs, fx) {
  if (is.null(fx$dsep)) {
    return(NULL)
  }

  x <- fx$dsep$x
  y <- fx$dsep$y
  z <- unlist(fx$dsep$z)

  calls <- list(
    caugi = bquote(caugi::d_separated(graphs$cg, .(x), .(y), .(z))),
    bnlearn = bquote(bnlearn::dsep(graphs$bn, .(x), .(y), .(z))),
    dagitty = bquote(dagitty::dseparated(graphs$dg, .(x), .(y), .(z))),
    pcalg = bquote(pcalg::dsep(.(x), .(y), .(z), graphs$gNEL))
  )

  run_bench(calls, "d_separated", fx)
}
```

```{r}
#| label: plot-dsep
#| purl: false

plot_op("d_separated")
```

`caugi` is the clear winner here, staying near a tenth of a millisecond on every
fixture, including the dense $n = 10000$ graph. `pgmpy` is the runner-up, within
a few milliseconds throughout. This is the one operation where Tetrad's compiled
code does *not* save it: its m-separation routine explores paths in a way that
scales badly on dense graphs, reaching over a second at $n = 1000$ and roughly
nine seconds at $n = 10000$---a striking example of an algorithmic choice
dominating the language advantage. Among the R packages, `bnlearn` and `pcalg`
pay to build a moralized graph on each call (`pcalg` into the hundreds of
milliseconds at $n = 1000$, slowest of the three R competitors), and `dagitty`
re-parses its DSL; all three are skipped at $n = 10000$.

### Subgraph Extraction

Subgraph extraction is where we explicitly test graph **building** performance.
For `caugi` we time `subgraph(cg, nodes) |> build()` together so the CSR rebuild
cost is captured. `igraph`, `bnlearn`, `pgmpy`, and Tetrad all return
adjacency-backed subgraphs without an analogous index rebuild step. `pcalg`
exposes no native subgraph constructor (users typically call `graph::subGraph()`
on the underlying `graphNEL`), so it is omitted here.

```{r}
#| label: bench-subgraph
#| eval: false
#| echo: true

bench_subgraph <- function(graphs, fx) {
  sub <- unlist(fx$subgraph_nodes)

  calls <- list(
    caugi = bquote({
      sg <- caugi::subgraph(graphs$cg, .(sub))
      caugi::build(sg)
    }),
    igraph = bquote(igraph::subgraph(graphs$ig, .(sub))),
    bnlearn = bquote(bnlearn::subgraph(graphs$bn, .(sub)))
  )

  run_bench(calls, "subgraph", fx)
}
```

```{r}
#| label: plot-subgraph
#| purl: false

plot_op("subgraph")
```

`caugi` and `igraph` are the only packages that stay in the low-single-digit
milliseconds at $n = 10000$ (around 1-2 ms). The two run neck-and-neck: `caugi`
is ahead at $n = 100$ and $n = 1000$, while `igraph` edges it out on the largest
fixture---remarkably close given that `caugi` pays for a full CSR rebuild on
every call. Everyone else is one to three
orders of magnitude slower on the largest fixtures, copying nodes and edges into
a fresh object each call: `pgmpy` reaches tens to hundreds of milliseconds,
Tetrad several hundred, and `bnlearn` over two seconds. The takeaway is that
even though `caugi` graphs are nominally expensive to construct, the rebuild
after `subgraph()` is fast enough to beat adjacency-list constructors that
rebuild no indexes at all.

## Summary

Across the seven operations, `caugi` either leads or sits in the leading group:

- It matches or beats `igraph` and `bnlearn` on the $O(1)$ adjacency lookups
  (`parents`, `children`, Markov blanket).
- It is the fastest R package by a wide margin on transitive-closure operations
  (`ancestors`, `descendants`, d-separation), where the CSR's precomputed
  forward and reverse adjacency pays off.
- Even on `subgraph()`, where the CSR has to be rebuilt, it stays in the
  low-single-digit milliseconds at $n = 10000$ (matched only by `igraph`), while
  `bnlearn`, `pgmpy`, and Tetrad reach hundreds of milliseconds to seconds.

The most striking pattern emerges at $n = 10000$: `caugi` is the only package
that stays fast across *every* operation. Tetrad (compiled Java) and `pgmpy`
(NetworkX-backed) beat `caugi` on the simplest lookups by a small constant
factor, which is unsurprising given the R-level dispatch overhead `caugi` pays
on every call. But several competitors either run out of room (`dagitty`, `ggm`,
and the moralization-based d-separation tests are too slow to include at
$n = 10000$) or collapse on specific operations. Tetrad in particular, despite
being compiled, takes seconds on dense descendants and d-separation queries
because of how those routines scale, not because of the language.

The headline holds: by frontloading work into the build step and querying a
CSR-backed structure, `caugi` keeps query times effectively constant in graph
size for the operations users hit most often. The price is paid up front, in the
rebuild step that runs lazily on the first query after a mutation.

## Reproducing These Benchmarks

The R benchmark code shown in the chunks above is the actual source: each chunk
has `eval = FALSE` so the rendered vignette does not run them, and the wrapper
script at
[`tools/benchmark/run_r_bench.R`](https://github.com/frederikfabriciusbjerre/caugi/tree/main/tools/benchmark/run_r_bench.R)
extracts them with `knitr::purl()` and sources the result into the R session.

The full harness lives in
[`tools/benchmark/`](https://github.com/frederikfabriciusbjerre/caugi/tree/main/tools/benchmark)
and uses [Task](https://taskfile.dev) to orchestrate the cross-language runners.
To reproduce just the R numbers (no Python or Java toolchains needed):

```sh
cd tools/benchmark
task fixtures && task r
```

`task r` purls the vignette into `tools/benchmark/bench_r.R`, runs it, prints a
per-package, per-operation summary, and writes the raw timings to
`results/r.csv`. Do not edit `bench_r.R` directly---change the chunks in this
vignette instead. See `tools/benchmark/README.md` for the full pipeline
(`task all`) and per-language details.

```{r}
#| label: bench-driver
#| include: false
#| eval: false

results <- list()

for (fx in spec$fixtures) {
  message(sprintf("[bench_r] %s (n=%d, edges=%d)", fx$id, fx$n, fx$n_edges))
  graphs <- load_fixture(fx)
  results[[length(results) + 1L]] <- bench_parents(graphs, fx)
  results[[length(results) + 1L]] <- bench_children(graphs, fx)
  results[[length(results) + 1L]] <- bench_ancestors(graphs, fx)
  results[[length(results) + 1L]] <- bench_descendants(graphs, fx)
  results[[length(results) + 1L]] <- bench_markov_blanket(graphs, fx)
  results[[length(results) + 1L]] <- bench_dsep(graphs, fx)
  results[[length(results) + 1L]] <- bench_subgraph(graphs, fx)
}

results <- results[!vapply(results, is.null, logical(1))]
out <- data.table::rbindlist(results)
data.table::fwrite(out, file = file.path(RESULTS_DIR, "r.csv"))

message(sprintf("[bench_r] wrote %d rows to %s/r.csv", nrow(out), RESULTS_DIR))

op_levels <- c(
  "parents",
  "children",
  "ancestors",
  "descendants",
  "markov_blanket",
  "d_separated",
  "subgraph"
)

summary_dt <- out[,
  .(median_ms = median(median_ns) / 1e6),
  by = .(package, operation)
]
summary_dt[, operation := factor(operation, levels = op_levels)]
summary_wide <- data.table::dcast(
  summary_dt,
  package ~ operation,
  value.var = "median_ms"
)

message(
  "\n[bench_r] median ms per (package, operation), aggregated over fixtures:"
)

print(summary_wide, digits = 3, row.names = FALSE)
```

## Session Info

```{r}
#| label: session-info
#| purl: false

sessionInfo()
```
